---
title: "Peer-graded Assignment: Course Project: Shiny Application and Reproducible Pitch"
author: "Haoyi Wei"
date: "2022-12-21"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Project Requirement
This peer assessed assignment has two parts.

- My Shiny Application
   - Write a shiny application with associated supporting documentation.This application uses data about the traffic fatalities in U.S between 1994 to 2020. The application provides visualization about traffic fatalities in states across the U.S. from 1994 to 2020.

- My Reproducible Pitch Presentation
   - I am going to create a web page using Slidify or Rstudio Presenter with an html5 slide deck.

## Data Source: Fatality Analysis Reporting System (FARS)

FARS is a nationwide census providing NHTSA, Congress and the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.

Data Link: <https://www-fars.nhtsa.dot.gov/states/statesfatalitiesfatalityrates.aspx>

The following calculations are done in the server :

- Load data.
- Reshape data.
- Subset data.

## Data and Server calculations

Summarize the data

```{r, comment=""}
library(reshape)
fatals <- read.csv("./Fatal.csv")
fatalities <- melt(fatals, id=c("State","ID"))
fatalities$variable <- as.numeric(gsub("X","",fatalities$variable))
str(fatalities)
```

## Shiny App
```{r, echo = FALSE}
knitr::include_app("https://econhaoyiwei.shinyapps.io/DataScience9_Assignment3/")
```