---
title: "Regression Models"
author: "Haoyi Wei"
date: "2022-12-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Peer-graded Assignment: regression Models Course Project

## Executive summary

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

+ “Is an automatic or manual transmission better for MPG”
+ "Quantify the MPG difference between automatic and manual transmissions"  

## Analysis

### Import the dataset

```{r}
# load the mtcars data
data(mtcars)

```

### Exploratory Analysis

```{r}
names(mtcars)
dim(mtcars)
str(mtcars)
summary(mtcars)
```

```{r}
library(explore)
library(dplyr)

# how many variables?
explore_tbl(mtcars)

# describe the dataset
describe(mtcars)

```

```{r, fig.width=14, fig.height=40}
# explore the variables
explore_all(mtcars)
```

### Descriptive Analysis

```{r}
# descriptive analysis:is an automatic or manual transmission better for MPG

library(ggplot2)
library(dplyr)

mtcars$am <- factor(mtcars$am)
levels(mtcars$am) <- list(Manual=1, Automatic=0)
ggplot(mtcars, aes(x=am, y=mpg)) +
          geom_boxplot(color="blue") +
          ggtitle("MPG by Transmission Type") +
          labs(x="Transmission Type", y="MPG")

```
The mean and median of MPG different between manual and auto

### Analytical Analysis


```{r}
data(mtcars)
model <- lm(mpg~., data=mtcars)
summary(model)
```
Detect [multicollinearity](https://www.codingprof.com/3-ways-to-test-for-multicollinearity-in-r-examples/)

```{r, fig.width=20, fig.height=20}
# Detect multicollinearity with correlation matrix
library("corrplot")
mtcars$am <- as.numeric(mtcars$am)
corrplot(cor(mtcars), method = "number")

# Test for Multicollinearity with Variance Inflation Factors (VIF)

#TOLERANCE & VARIANCE INFLATION FACTOR (VIF)
library("olsrr")
ols_vif_tol(model)
```


As a rule of thumb, a VIF exceeding 5 requires further investigation, whereas VIFs above 10 indicate multicollinearity. Ideally, the Variance Inflation Factors are below 3. The result indicate the possibilities of multicollinearity

To address the multicollinearity issue, we use the stepwise selection method
```{r}
library(MASS)
step <- stepAIC(model, direction="both", trace=FALSE)
summary(step)$coeff
```


The p-value for am is greater than 0.1, we can not conclude the coefficient on am is different from zero at convention significant levels.

```{r}
# fit the new model
final_model <- lm(mpg ~ wt+qsec+factor(am), data = mtcars)
summary(final_model)$coef
```

### Conclusion

On average, manual transmission cars have 2.94 MPGs more than automatic transmission cars

### Appendix

Residual plot

```{r}

final.res = resid(final_model)

# We now plot the residual against the observed values of the variable waiting.
#produce residual vs. fitted plot
plot(fitted(model), final.res)

```

The distribution of the residul fairly consitent across different level of fitted level. No major concern about the model design



```{r}
plot(final_model)
```

