---
title: "CodeBook"
author: "Haoyi Wei"
date: "2022-12-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=F)
```

# You should create one R script called run_analysis.R that does the following.

#read the datasets

```{r}
features <- read.table("~/Documents/GitHub/Coursera_Data_Science_Certificate/3_Data_cleaning/UCI HAR Dataset/features.txt", col.names = c("n","functions"))
activities <- read.table("~/Documents/GitHub/Coursera_Data_Science_Certificate/3_Data_cleaning/UCI HAR Dataset/activity_labels.txt", col.names = c("code", "activity"))
subject_test <- read.table("~/Documents/GitHub/Coursera_Data_Science_Certificate/3_Data_cleaning/UCI HAR Dataset/test/subject_test.txt", col.names = "subject")
x_test <- read.table("~/Documents/GitHub/Coursera_Data_Science_Certificate/3_Data_cleaning/UCI HAR Dataset/test/X_test.txt", col.names = features$functions)
y_test <- read.table("~/Documents/GitHub/Coursera_Data_Science_Certificate/3_Data_cleaning/UCI HAR Dataset/test/y_test.txt", col.names = "code")
subject_train <- read.table("~/Documents/GitHub/Coursera_Data_Science_Certificate/3_Data_cleaning/UCI HAR Dataset/train/subject_train.txt", col.names = "subject")
x_train <- read.table("~/Documents/GitHub/Coursera_Data_Science_Certificate/3_Data_cleaning/UCI HAR Dataset/train/X_train.txt", col.names = features$functions)
y_train <- read.table("~/Documents/GitHub/Coursera_Data_Science_Certificate/3_Data_cleaning/UCI HAR Dataset/train/y_train.txt", col.names = "code")

```

# Merges the training and the test sets to create one data set.

```{r}
X <- rbind(x_train, x_test)
Y <- rbind(y_train, y_test)
Subject <- rbind(subject_train, subject_test)
Rawdata <- cbind(Subject, Y, X)
```

# Extracts only the measurements on the mean and standard deviation for each measurement.

```{r}
extracts <- Rawdata %>% 
          select(subject, code, contains("mean"), contains("std"))
```

# Uses descriptive activity names to name the activities in the data set

```{r}
extracts$code <- activities[extracts$code, 2]
```

# Appropriately labels the data set with descriptive variable names.

```{r}
names(extracts)[2] = "activity"
names(extracts)<-gsub("Acc", "Accelerometer", names(extracts))
names(extracts)<-gsub("Gyro", "Gyroscope", names(extracts))
names(extracts)<-gsub("BodyBody", "Body", names(extracts))
names(extracts)<-gsub("Mag", "Magnitude", names(extracts))
names(extracts)<-gsub("^t", "Time", names(extracts))
names(extracts)<-gsub("^f", "Frequency", names(extracts))
names(extracts)<-gsub("tBody", "TimeBody", names(extracts))
names(extracts)<-gsub("-mean()", "Mean", names(extracts), ignore.case = TRUE)
names(extracts)<-gsub("-std()", "STD", names(extracts), ignore.case = TRUE)
names(extracts)<-gsub("-freq()", "Frequency", names(extracts), ignore.case = TRUE)
names(extracts)<-gsub("angle", "Angle", names(extracts))
names(extracts)<-gsub("gravity", "Gravity", names(extracts))
```

# From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r}
install.packages("collapse")

secdata <- extracts %>%
          group_by(subject, activity) %>%
          summarise_all(mean)

write.table(secdata, "~/Documents/GitHub/Coursera_Data_Science_Certificate/3_Data_cleaning/secdata.txt", row.name=FALSE)

```
