---
title: "Course4 Exploratory Data Analysis - Assignment 1"
author: "Haoyi Wei"
date: "2022-12-22"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Programming Assignment 1

# Instrution

This assignment uses data from the UC Irvine Machine Learning Repository, a popular repository for machine learning datasets. In particular, we will be using the “Individual household electric power consumption Data Set” which I have made available on the course web site:

- Dataset: Electric power consumption [20Mb]
- Description: Measurements of electric power consumption in one household with a one-minute sampling rate over a period of almost 4 years. Different electrical quantities and some sub-metering values are available.
The following descriptions of the 9 variables in the dataset are taken from the UCI web site:

- Date: Date in format dd/mm/yyyy
- Time: time in format hh:mm:ss
- - Global_active_power: household global minute-averaged active power (in kilowatt)
- Global_reactive_power: household global minute-averaged reactive power (in kilowatt)
- Voltage: minute-averaged voltage (in volt)
- Global_intensity: household global minute-averaged current intensity (in ampere)
- Sub_metering_1: energy sub-metering No. 1 (in watt-hour of active energy). It corresponds to the kitchen, containing mainly a dishwasher, an oven and a microwave (hot plates are not electric but gas powered).
- Sub_metering_2: energy sub-metering No. 2 (in watt-hour of active energy). It corresponds to the laundry room, containing a washing-machine, a tumble-drier, a refrigerator and a light.
- Sub_metering_3: energy sub-metering No. 3 (in watt-hour of active energy). It corresponds to an electric water-heater and an air-conditioner.

Goal: reconstruct four plots

# Review criteria

- Was a valid GitHub URL containing a git repository submitted?
- Does the GitHub repository contain at least one commit beyond the original fork?
- Please examine the plot files in the GitHub repository. Do the plot files appear to be of the correct graphics file format?
- Does each plot appear correct?
- Does each set of R code appear to create the reference plot?


# Get and clean the data

```{r}
library(dplyr)

## Import the data
data0 <- read.table("~/Documents/GitHub/Coursera_Data_Science_Certificate/4_Exploratory Data Analysis/Code/household_power_consumption.txt", header=TRUE, sep=";", na.strings = "?", colClasses = c('character','character','numeric','numeric','numeric','numeric','numeric','numeric','numeric'))

# clean the data
data <- data0 %>%
          mutate(Date=as.Date(Date, "%d/%m/%Y")) %>% ## Format date to Type Date
          filter(Date>="2007-02-01" & Date<="2007-02-02") %>% ## Filter data set from Feb. 1, 2007 to Feb. 2, 2007
          na.omit() %>% ## Remove incomplete observation
          mutate(datetime=as.POSIXct(paste(Date,Time),format="%Y-%m-%d %H:%M:%S")) %>% #add time date column
          select(-c(Date,Time)) %>%
          relocate(datetime)
```
# Plot 1
```{r}
## plot 1
## Create the histogram
hist(data$Global_active_power, main="Global Active Power", xlab = "Global Active Power (kilowatts)", col="red")
```

# Plot 2
```{r}
## plot 2
## Create Plot 2
plot(data$Global_active_power~data$datetime, type="l", ylab="Global Active Power (kilowatts)", xlab="Datetime", main="Global Active Power by Time of the Day")
```

# Plot 3
```{r}
with(data, {
          plot(Sub_metering_1~datetime, type="l",
               ylab="Energy Sub Metering", xlab="Datetime", main="Global Active Power by Time of the Day and Meter Type")
          lines(Sub_metering_2~datetime,col='Red')
          lines(Sub_metering_3~datetime,col='Blue')
})
legend("topright", col=c("black", "red", "blue"), lwd=c(1,1,1), 
       c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"))

```

# Plot 4
```{r}
# par can be used to set graphical parameters
par(mfrow=c(2,2), mar=c(4,4,2,1), oma=c(0,0,2,0))
with(data, {
          plot(Global_active_power~datetime, type="l", 
               ylab="Global Active Power", xlab="Datetime")
          plot(Voltage~datetime, type="l", 
               ylab="Voltage", xlab="Datetime")
          plot(Sub_metering_1~datetime, type="l", 
               ylab="Global Sub Metering)", xlab="Datetime")
          lines(Sub_metering_2~datetime,col='Red')
          lines(Sub_metering_3~datetime,col='Blue')
          legend("topright", col=c("black", "red", "blue"), lty=1, lwd=2, bty="n",
                 legend=c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"))
          plot(Global_reactive_power~datetime, type="l", 
               ylab="Global Rective Power",xlab="Datetime")
})
```


