---
title: "Assignment1_part2"
author: "Haoyi Wei"
date: "2022-12-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Analyze the ToothGrowth data in the R datasets package.

## Overview
This project analyze the ToothGrowth data in the R datasets package.

## Load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
data(ToothGrowth)
```

## Provide a basic summary of the data

```{r}
head(ToothGrowth)
dim(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)

library(ggplot2)
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
ggplot(ToothGrowth, aes(x=dose, y= len, fill=supp)) +
          geom_boxplot() +
          scale_fill_manual(values=c("yellow","orange"))
```


### 1. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
```{r}

library(dplyr)
# Hypothesis Testing
dose.5 <- ToothGrowth %>%
          filter(dose ==0.5)
dose1 <- ToothGrowth %>%
          filter(dose ==1)
dose2 <- ToothGrowth %>%
          filter(dose ==2)
```

#### t-test for 0.5 mg/day dose
```{r}
t.test(len ~ supp, dose.5)
```

#### t-test for 1 mg/day dose
```{r}
t.test(len ~ supp, dose1)
```

#### t-test for 2 mg/day dose
```{r}
t.test(len ~ supp, dose2)
```

#### Summarize the results


Summarize the results in a table
```{r}
dose <- c(0.5, 1.0, 2.0)
p_value <- c(0.0064, 0.0010, 0.9639)
cv <- c("1.72, 8.78", "2.80, 9.06", "-3.80, 3.64")
outcome <- c("Reject null", "Reject null", "Fail to reject null")
data.frame(dose, cv, p_value, outcome)
```
For dose 0.5 and dose 1, the p value are indistinguishable from zero, thus reject the null hypotheses that the supplement types don’t have a difference on tooth growth. 

For dose 2.0, fail to reject the null because the p-value is greater than 0.5.

### Conclusion

For low dose such as 0.5 and 1, OJ is higher than VC, but for dose 2, OJ and VC's effect is inditingusihable.